---
title: "BayesCapstone"
output: html_document
---

```{r}
library(dplyr)
library(rjags)
library(ggplot2)
```


#=======================DATA_CLEANING
```{r}
weatherdf <- read.csv("weatherAUS.csv")
weatherdf <- weatherdf%>%select(c(Date, Location, MinTemp, MaxTemp, RainToday, RainTomorrow))

#  2107 rows of missing data
dim(weatherdf %>% filter(is.na(MinTemp) | is.na(MaxTemp) | is.na(RainToday) | is.na(RainTomorrow)))


# split Date to year-month-day, filter out 2107 rows because of missing data
library(tidyr)
library(lubridate)
weatherdf <- weatherdf %>% mutate(year=year(Date), month=month(Date), day=day(Date)) %>% drop_na(MinTemp, MaxTemp, RainToday, RainTomorrow)

Queensland = c("Brisbane", "Cairns", "GoldCoast", "Townsville" )
New_South_Wales = c("Albury", "BadgerysCreek", "Cobar" , "Moree", "Newcastle", "NorahHead", "Penrith", "Sydney", "SydneyAirport", "WaggaWagga", "Williamtown", "Wollongong", "Canberra", "Tuggeranong", "MountGinini", "CoffsHarbour")
Victoria = c("Richmond", "Ballarat", "Bendigo", "Sale", "MelbourneAirport", "Melbourne", "Mildura", "Nhil", "Portland", "Watsonia", "Dartmoor")
Tasmania = c("Launceston", "Hobart")
South_Australia = c("Woomera", "Adelaide", "MountGambier", "Nuriootpa" )
Western_Australia = c("Walpole", "PearceRAAF", "PerthAirport", "Perth", "SalmonGums", "Albany", "Witchcliffe")
Northern_Australia = c("Uluru", "Katherine", "Darwin", "AliceSprings")
"NorfolkIsland"


weatherdf <- weatherdf %>% mutate(state=(ifelse(Location %in% Queensland, "Queensland", ifelse(Location %in% New_South_Wales, "New_South_Wales", ifelse(Location %in% Victoria, "Victoria", ifelse(Location %in% Tasmania, "Tasmania", ifelse(Location %in% South_Australia, "South_Australia", ifelse(Location %in% Western_Australia, "Western_Australia", ifelse(Location %in% Northern_Australia, "Northern_Australia", ifelse(Location=="NorfolkIsland", "NorfolkIsland", NA))))))))))


weatherdf <- weatherdf %>% mutate(RainTomorrow=(ifelse(RainTomorrow=="Yes", 1,0))) %>% mutate(TempDiff= (MaxTemp-MinTemp))



# split data into train and test, used when we have the model
## 75% of the sample size
smp_size <- floor(0.75 * nrow(weatherdf))
## set the seed to make your partition reproducible
set.seed(123)
train_ind <- sample(seq_len(nrow(weatherdf)), size = smp_size)
train <- weatherdf[train_ind, ]
test <- weatherdf[-train_ind, ]
dim(train)
dim(test)


# run data is used when test whether the code works or not. For shorten the rjags running time.
# only contains 5% of original data
runsample <- floor(0.05 * nrow(weatherdf))
set.seed(123)
run_ind <- sample(seq_len(nrow(weatherdf)), size = runsample)
run <- weatherdf[run_ind,]

```



```{r}
options(digits=6)
# four plots: 
#Model 1: RainTmr~1
ggplot(weatherdf, aes(x=factor(RainTomorrow))) + geom_bar(alpha=0.5)

#Model 2: RainTmr~TemDiff 
ggplot(weatherdf,aes(weatherdf$TempDiff, fill = factor(RainTomorrow))) +
  geom_density(alpha = 0.8) 

#Model 3: RainTmr~ State+ TemDiff 
ggplot(weatherdf, aes(x= weatherdf$TempDiff, fill=factor(RainTomorrow))) + geom_density(alpha = 0.5)+facet_wrap(~state)

#To get a sense of TempDiff in each state: TempDiff~state
ggplot(weatherdf, aes(x= weatherdf$TempDiff, fill=state)) + geom_density(alpha = 0.5)
```




#=================================MODELS
Y: rain tomorrow or not (in [0,1])
X: tempreture difference
$\theta$: overall tomorrow raining rate
i: different states
j: whether rain today(in [0,1])

**Model 1** 
Whether it rains tomorrow depends on the overall raining rate
$$
\begin{split}  
Y_{i} | \theta, b_0& \sim  Bern(\theta) \\ 
log(\frac{\theta}{1-\theta})  & = b_0 \\
b_0 & \sim N(0.3, \frac{1}{2500})
\end{split}
$$


**Model 2**
Whether it rains tomorrow depends on the overall raining rate, which changes in different tempreture difference
$$
\begin{split}  
Y_{i} | \theta, b_0,b_1 & \sim  Bern(\theta) \\ 
log(\frac{\theta}{1-\theta})  & = b_0+b_1*X_{i} \\
b_0,b_1 & \sim N(0.3, \frac{1}{2500})
\end{split}
$$


**Model3**
Whether it rains tomorrow depends on the raining rate in specific state, each of which has a different trend with tempreture difference. The individual states’s intercept and slope are nornally distributed around global intercept and slope. 

$$
\begin{split}  
Y_{ij} | \theta, b_{0},b_{1},\beta_0,\beta_1, \tau_0, \tau_1 & \sim  Bern(\theta_i) \\ 
log(\frac{\theta_i}{1-\theta_i})  & = b_{0i}+b_{1i}*X_{ij} \\
b_{0i}|\beta_0, \tau_0 & \sim N(\beta_0,\tau_0^{-1}) \\
b_{1i}|\beta_1, \tau_1 & \sim N(\beta_1,\tau_1^{-1}) \\
\beta_0, \beta_1 &\sim N(0, \frac{1}{10000}) \\
\tau_0, \tau_1 &\sim Gamma(7, 10000)
\end{split}
$$



```{r cache=TRUE}
# DEFINE RAIN MODEL 1
rain_model_1 <- "model{  
    # Data
    for(i in 1:length(y)) {
        y[i] ~ dbern(theta)
    }
    logit(theta) <- b0

    # Prior
    b0   ~ dnorm(m0, t0)
}"


# COMPILE
rain_jags_1 <- jags.model(textConnection(rain_model_1), 
    data = list(y = train$RainTomorrow, m0=0.3, t0=1/2500),
    inits=list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 454))

rain_sim_1 <- coda.samples(rain_jags_1,
    variable.names = c("b0"),
    n.iter = 10000)

rain_chains_1 <- data.frame(iteration = 1:10000, rain_sim_1[[1]])
plot(rain_sim_1)
summary(rain_sim_1)
```



```{r}
# TEST-MODEL 1
rain_chains_1 <- rain_chains_1 %>% 
        mutate(theta_pred = exp(b0) / (1+exp(b0)))

set.seed(454) # where should I put set.seed? If put it into the function, the probability for each row will be the same
my_predict <- function(x){
    predict_chains <- rain_chains_1 %>% 
        mutate(y_prediction = rbinom(nrow(rain_chains_1), size=1, prob=theta_pred))
    
    mean(predict_chains$y_prediction)
}


test <- test %>% 
    mutate(prob_pred1 = Vectorize(my_predict)(x = row_number()))

tab <- table(test$RainTomorrow, (test$prob_pred1 >= 0.2))
tab
prop.table(tab, margin = 1)
sum(diag(tab)) / sum(tab)

   
#     FALSE
#   0 27178
#   1  7844
#    
#     FALSE
#   0     1
#   1     1
# [1] 0.776026
```


```{r cache=TRUE}
# DEFINE RAIN MODEL 2
rain_model_2 <- "model{  
    # Data
    for(i in 1:length(y)) {
        y[i] ~ dbern(theta[i])
        logit(theta[i]) <- b0 + b1 * x[i]
}   
        

    # Priors
    b0   ~ dnorm(m0, t0)
    b1   ~ dnorm(m1, t1)
}"

# COMPILE
rain_jags_2 <- jags.model(textConnection(rain_model_2), 
    data = list(y = train$RainTomorrow, x = train$TempDiff, m0 = 0.3, m1 = 0.3, t0 = 1/2500, t1 = 1/2500),
    inits=list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 454))

rain_sim_2 <- coda.samples(rain_jags_2,
    variable.names = c("b0", "b1"),
    n.iter = 10000)

rain_chains_2 <- data.frame(rain_sim_2[[1]])
plot(rain_sim_2)
summary(rain_sim_2)
```

```{r}
# TEST-MODEL 2
my_predict <- function(x){ 
  # input is TempDiff 
  set.seed(454)
  predict_chains <- rain_chains_2 %>% 
      mutate(theta_pred = exp(b0+b1*x) / (1+exp(b0+b1*x))) %>%  
      mutate(y_prediction = rbinom(nrow(rain_chains_2), size=1, prob=theta_pred))
    
    mean(predict_chains$y_prediction)
}

test <- test %>% 
    mutate(prob_pred2 = Vectorize(my_predict)(x = test$TempDiff))

tab <- table(test$RainTomorrow, (test$prob_pred2 >= 0.2))
tab
prop.table(tab, margin = 1)
sum(diag(tab)) / sum(tab)

#     FALSE  TRUE
#   0 26663   515
#   1  6768  1076
#    
#         FALSE      TRUE
#   0 0.9810508 0.0189492
#   1 0.8628251 0.1371749
# [1] 0.792045
```







```{r cache=TRUE}
# DEFINE RAIN MODEL 3
rain_model_3 <- "model{  
    # Data
    for(i in 1:length(y)) {
        y[i] ~ dbern(theta[i])
        logit(theta[i]) <- b0[subject[i]] + b1[subject[i]] * x[i]
    }

    # Data: subjects
    for(i in 1:n_subjects) {
        b0[i] ~ dnorm(beta0, tau0)
        b1[i] ~ dnorm(beta1, tau1)
    }

    # Priors
    beta0 ~ dnorm(0, 1/10000)
    beta1 ~ dnorm(0, 1/10000)
    tau0 ~ dgamma(7, 10000)
    tau1 ~ dgamma(7, 10000)
}"

# COMPILE
rain_jags_3 <- jags.model(textConnection(rain_model_3), 
    data = list(y = train$RainTomorrow, x = train$TempDiff, subject = factor(train$state), n_subjects = 8),
    inits=list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 454))

rain_sim_3 <- coda.samples(rain_jags_3,
    variable.names = c("beta0", "beta1", "tau0", "tau1", "b0", "b1"),
    n.iter = 10000)

rain_chains_3 <- data.frame(rain_sim_3[[1]])

plot(rain_sim_3)
summary(rain_sim_3)
```




```{r}
# TEST-MODEL 3
my_predict <- function(state, x){ 
  # input is TempDiff and state
  b0 <- rain_chains_3[,state]
  b1 <- rain_chains_3[,state+8]
  
  set.seed(454)
  predict_chains <- rain_chains_3 %>% 
      mutate(theta_pred = exp(b0+b1*x) / (1+exp(b0+b1*x))) %>%  
      mutate(y_prediction = rbinom(nrow(rain_chains_3), size=1, prob=theta_pred))
    
    mean(predict_chains$y_prediction)
}


# what does b0.1. and b1.1. correspond to? In alphabetic order? 
stateSet <- sort(unique(test$state))
state_index <- match(test$state,stateSet)


test <- test %>% 
    mutate(prob_pred3 = Vectorize(my_predict)(state=state_index, x = test$TempDiff))

tab <- table(test$RainTomorrow, (test$prob_pred3 >= 0.2))
tab
prop.table(tab, margin = 1)
sum(diag(tab)) / sum(tab)

#     FALSE  TRUE
#   0 26575   603
#   1  6587  1257
#    
#         FALSE      TRUE
#   0 0.9778129 0.0221871
#   1 0.8397501 0.1602499
# [1] 0.7947

```



```{r cache=TRUE}
# DEFINE RAIN MODEL 4--location(city) instead of state
rain_model_4 <- "model{  
    # Data
    for(i in 1:length(y)) {
        y[i] ~ dbern(theta[i])
        logit(theta[i]) <- b0[subject[i]] + b1[subject[i]] * x[i]
    }

    # Data: subjects
    for(i in 1:n_subjects) {
        b0[i] ~ dnorm(beta0, tau0)
        b1[i] ~ dnorm(beta1, tau1)
    }

    # Priors
    beta0 ~ dnorm(0, 1/10000)
    beta1 ~ dnorm(0, 1/10000)
    tau0 ~ dgamma(7, 10000)
    tau1 ~ dgamma(7, 10000)
}"

# COMPILE
rain_jags_4 <- jags.model(textConnection(rain_model_4), 
    data = list(y = train$RainTomorrow, x = train$TempDiff, subject = factor(train$Location), n_subjects = 49),
    inits=list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 454))

rain_sim_4 <- coda.samples(rain_jags_4,
    variable.names = c("beta0", "beta1", "tau0", "tau1", "b0", "b1"),
    n.iter = 10000)

rain_chains_4 <- data.frame(rain_sim_4[[1]])

plot(rain_sim_4)
summary(rain_sim_4)
```


```{r}
# TEST-MODEL 4
my_predict <- function(location, x){ 
  # input is TempDiff and state
  b0 <- rain_chains_4[,location]
  b1 <- rain_chains_4[,location+49]
  
  set.seed(454)
  predict_chains <- rain_chains_4 %>% 
      mutate(theta_pred = exp(b0+b1*x) / (1+exp(b0+b1*x))) %>%  
      mutate(y_prediction = rbinom(nrow(rain_chains_4), size=1, prob=theta_pred))
    
    mean(predict_chains$y_prediction)
}


# In alphabetic order? 
locationSet <- sort(unique(test$Location))
location_index <- match(test$Location,locationSet)


test <- test %>% 
    mutate(prob_pred4 = Vectorize(my_predict)(location=location_index, x = test$TempDiff))

tab <- table(test$RainTomorrow, (test$prob_pred4 >= 0.2))
tab
prop.table(tab, margin = 1)
sum(diag(tab)) / sum(tab)

# export dataset
write.csv(test, "test.csv")
write.csv(rain_chains_4, "rain_chains_4.csv")
```




#=============================================SHINNY APP for the fourth model

```{r echo=FALSE}
library(dplyr)
new_predict <- function(location, TempDiff){ 
  locationSet <- sort(unique(test$Location))
  location <- match(location,locationSet)
  
  # input is TempDiff and state
  b0 <- rain_chains_4[,location]
  b1 <- rain_chains_4[,location+49]
  
  set.seed(454)
  predict_chains <- rain_chains_4 %>% 
      mutate(theta_pred = exp(b0+b1*TempDiff) / (1+exp(b0+b1*TempDiff))) %>%  
      mutate(y_prediction = rbinom(nrow(rain_chains_4), size=1, prob=theta_pred))
    
    return(mean(predict_chains$y_prediction))
}

# specific_accurary <- function(location,tempDiff, cut){
#   sub <- test %>% filter(Location==location) %>% filter(TempDiff==tempDiff)
#   tab <- table(sub$RainTomorrow, (sub$prob_pred4 >= cut))
#   confMatrix <- as.data.frame.matrix(prop.table(tab, margin = 1))
#   confMatrix <- cbind(model = c("not rain","rain"), confMatrix)
#   return(confMatrix)
# }

model_accurary <- function(cut){
  tab <- table(test$RainTomorrow, (test$prob_pred4 >= cut))
  confMatrix <- prop.table(tab, margin = 1)
  confMatrix <- as.data.frame.matrix(confMatrix)
  confMatrix <- cbind(model = c("not rain","rain"), confMatrix)
  acc <- sum(diag(tab)) / sum(tab)
  return(list(Matrix=confMatrix, Acc=acc))
}

rain_predict <- function(location, TempDiff, cut){
  prob <- new_predict(location, TempDiff)
  result=NULL
  if (prob>=cut){result="rain"}
  else {result="not rain"}
  return(result)
}


library(shiny)
city = sort(unique(train$Location))
server <- function(input, output) {
  output$modelTable <- renderTable({
      model_accurary(input$cut)$Matrix
    })
  # output$specificTable <- renderTable({
  #       specific_accurary(input$city, input$TempDiff, input$cut)
  #   })
  output$acc <- renderText(
    model_accurary(input$cut)$Acc
  )
  output$prediction <- renderText(
    new_predict(input$city, input$TempDiff)
  )
  output$rain <- renderText(
    rain_predict(input$city, input$TempDiff, input$cut)
  )
  
}


#build the user interface
ui <- fluidPage(
  titlePanel("Australia Weather Predictor"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("cut", "Threshold of Probability", min = 0, max = 1, value = 0.2, step=0.05),
      selectInput("city", "Location", city),
      sliderInput("TempDiff", "Temperature Difference of Present Day", min = 0, max = 50, value = 8)
    ),
    mainPanel(
      h3("***************CHOOSE THRESHOLD***************"),
      h4("The accurary of the model in this specific threshold"), 
      tableOutput("modelTable") ,
      h4("The overall accuracy of the model is: "), 
      textOutput("acc"),
      # h4("The accurary of the model in this specific location and temperature difference"), 
      # tableOutput("specificTable") ,
      br(), br(),
      h3("*****************MAKE PREDICTION*****************"),
      h4("The probability to rain tomorrow is: "), 
      textOutput("prediction"),
      h4("Whether it will rain tomorrow: "), 
      textOutput("rain")
    )
  )
)


#check out the shiny app!
shinyApp(ui = ui, server = server)
```

```{r}
# Model Comparison Code
library(ggplot2)
results<-data.frame("Model"=c(1,1,1,2,2,2,3,3,3,4,4,4), "Values"=c(0, 1, 0.23, 0.59, 0.77, 0.63, 0.60, 0.77, 0.64, 0.61, 0.77, 0.65), "Type"=c("Specificity", "Sensitivity", "Overall.Accuracy","Specificity", "Sensitivity", "Overall.Accuracy","Specificity", "Sensitivity", "Overall.Accuracy","Specificity", "Sensitivity", "Overall.Accuracy"))
results

p <-ggplot(results, aes(Model, Values))
p +geom_bar(stat = "identity", aes(fill = Type), position = "dodge")
```


















